---
title: "02 - Data Merging"
author: "Murray"
date: "5/9/2020"
output: html_document
---

In 01, we retrieved the data and cleaned it up to some extent. But now we need to merge the data into one. 

This mainly involves fixing discrepencies in country names

```{r, include=F}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1 - Find the country name discrepencies

For now, we're just merging age data, obesity data, gdp data, and tests_cases_deaths data

```{r}
length(unique(tests_cases_deaths_data$location))
length(unique(obesity_data$Country))
length(unique(median_age_data$Country))
length(unique(gdp_per_capita_ppp_data$Country))

tests_cases_deaths_data$location <- gsub(" ", "", tests_cases_deaths_data$location)
obesity_data$Country <- gsub(" ", "", obesity_data$Country)
gdp_per_capita_ppp_data$Country <- gsub(" ", "", gdp_per_capita_ppp_data$Country)

# Make obesity_data names correspond to tests_cases_deaths_data names

setdiff(obesity_data$Country, tests_cases_deaths_data$location)
setdiff(tests_cases_deaths_data$location, obesity_data$Country)

obesity_data$Country[obesity_data$Country == "Bolivia(PlurinationalStateof)"] <- "Bolivia"
obesity_data$Country[obesity_data$Country == "Venezuela(BolivarianRepublicof)"] <- "Venezuela"
obesity_data$Country[obesity_data$Country == "UnitedKingdomofGreatBritainandNorthernIreland"] <- "UnitedKingdom"
obesity_data$Country[obesity_data$Country == "UnitedStatesofAmerica"] <- "UnitedStates"
obesity_data$Country[obesity_data$Country == "DemocraticRepublicoftheCongo"] <- "DemocraticRepublicofCongo"
obesity_data$Country[obesity_data$Country == "Côted'Ivoire"] <- "Coted'Ivoire"
obesity_data$Country[obesity_data$Country == "BruneiDarussalam"] <- "Brunei"
obesity_data$Country[obesity_data$Country == "CaboVerde"] <- "CapeVerde"
obesity_data$Country[obesity_data$Country == "Czechia"] <- "CzechRepublic"
obesity_data$Country[obesity_data$Country == "DemocraticPeople'sRepublicofKorea"] <- "SouthKorea"
obesity_data$Country[obesity_data$Country == "Eswatini"] <- "Swaziland"
obesity_data$Country[obesity_data$Country == "Iran(IslamicRepublicof)"] <- "Iran"
obesity_data$Country[obesity_data$Country == "LaoPeople'sDemocraticRepublic"] <- "Laos"
obesity_data$Country[obesity_data$Country == "RepublicofMoldova"] <- "Moldova"
obesity_data$Country[obesity_data$Country == "RepublicofNorthMacedonia"] <- "Macedonia"
obesity_data$Country[obesity_data$Country == "RussianFederation"] <- "Russia"
obesity_data$Country[obesity_data$Country == "SyrianArabRepublic"] <- "Syria"
obesity_data$Country[obesity_data$Country == "Timor-Leste"] <- "Timor"
obesity_data$Country[obesity_data$Country == "UnitedRepublicofTanzania"] <- "Tanzania"
obesity_data$Country[obesity_data$Country == "VietNam"] <- "Vietnam"
obesity_data$Country[obesity_data$Country == "RepublicofKorea"] <- "NorthKorea"

simple_model_data <- full_join(tests_cases_deaths_data %>% select(location, date,
                                                                  total_cases, total_cases_per_million,
                                                                  total_deaths, total_deaths_per_million) %>% rename(Country = location),
                               obesity_data %>% select(Country, Bothsexes) %>% rename(Obesity_rate = Bothsexes))
```

```{r}
setdiff(simple_model_data$Country, median_age_data$Country)[order(setdiff(simple_model_data$Country, median_age_data$Country))]
setdiff(median_age_data$Country, simple_model_data$Country)[order(setdiff(median_age_data$Country, simple_model_data$Country))]

simple_model_data$Country[simple_model_data$Country == "FaeroeIslands"] <- "FaroeIslands"
simple_model_data$Country[simple_model_data$Country == "UnitedStatesVirginIslands"] <- "VirginIslands"
simple_model_data$Country[simple_model_data$Country == "Micronesia(FederatedStatesof)"] <- "Micronesia"

median_age_data$Country[median_age_data$Country == "Bahamas,The"] <- "Bahamas"
median_age_data$Country[median_age_data$Country == "CaboVerde"] <- "CapeVerde"
median_age_data$Country[median_age_data$Country == "Congo,DemocraticRepublicofthe"] <- "DemocraticRepublicofCongo"
median_age_data$Country[median_age_data$Country == "Congo,Republicofthe"] <- "Congo"
median_age_data$Country[median_age_data$Country == "Czechia"] <- "CzechRepublic"
median_age_data$Country[median_age_data$Country == "Eswatini"] <- "Swaziland"
median_age_data$Country[median_age_data$Country == "Gambia,The"] <- "Gambia"
median_age_data$Country[median_age_data$Country == "Korea,North"] <- "NorthKorea"
median_age_data$Country[median_age_data$Country == "Korea,South"] <- "SouthKorea"
median_age_data$Country[median_age_data$Country == "Micronesia,FederatedStatesof"] <- "Micronesia"
median_age_data$Country[median_age_data$Country == "Timor-Leste"] <- "Timor"

simple_model_data <- full_join(simple_model_data, median_age_data)
```

```{r}
setdiff(simple_model_data$Country, gdp_per_capita_ppp_data$Country)[order(setdiff(simple_model_data$Country, gdp_per_capita_ppp_data$Country))]
setdiff(gdp_per_capita_ppp_data$Country, simple_model_data$Country)[order(setdiff(gdp_per_capita_ppp_data$Country, simple_model_data$Country))]

gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "BruneiDarussalam"] <- "Brunei"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "CaboVerde"] <- "CapeVerde"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "C\U3e34663cted'Ivoire"] <- "Coted'Ivoire"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "Eswatini"] <- "Swaziland"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "HongKongSAR"] <- "HongKong"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "IslamicRepublicofIran"] <- "Iran"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "Korea"] <- "SouthKorea"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "KyrgyzRepublic"] <- "Kyrgyzstan"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "LaoP.D.R."] <- "Laos"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "MacaoSAR"] <- "Macau"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "NorthMacedonia"] <- "Macedonia"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "RepublicofCongo"] <- "Congo"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "SlovakRepublic"] <- "Slovakia"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "St.KittsandNevis"] <- "SaintKittsandNevis"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "St.Lucia"] <- "SaintLucia"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "St.VincentandtheGrenadines"] <- "SaintVincentandtheGrenadines"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "TaiwanProvinceofChina"] <- "Taiwan"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "TheBahamas"] <- "Bahamas"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "TheGambia"] <- "Gambia"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "S\U3e33653coTom\U3e39653candPr\U3e64653cncipe"] <- "SaoTomeAndPrincipe"
gdp_per_capita_ppp_data$Country[gdp_per_capita_ppp_data$Country == "Timor-Leste"] <- "Timor"

simple_model_data <- full_join(simple_model_data, gdp_per_capita_ppp_data)
```

# Add in age distribution index

```{r}
setdiff(simple_model_data$Country, population_fatality_rates_relative_to_Italy$Country)[order(setdiff(simple_model_data$Country, population_fatality_rates_relative_to_Italy$Country))]
setdiff(population_fatality_rates_relative_to_Italy$Country, simple_model_data$Country)[order(setdiff(population_fatality_rates_relative_to_Italy$Country, simple_model_data$Country))]

risk_adjusted_age$Country[risk_adjusted_age$Country == "BruneiDarussalam"] <- "Brunei"
risk_adjusted_age$Country[risk_adjusted_age$Country == "CaboVerde"] <- "CapeVerde"
risk_adjusted_age$Country[risk_adjusted_age$Country == "ChinaMacaoSAR"] <- "Macau"
risk_adjusted_age$Country[risk_adjusted_age$Country == "ChinaHongKongSAR"] <- "HongKong"
risk_adjusted_age$Country[risk_adjusted_age$Country == "DemPeoplesRepublicofKorea"] <- "NorthKorea"
risk_adjusted_age$Country[risk_adjusted_age$Country == "DRCongo"] <- "DemocraticRepublicofCongo"
risk_adjusted_age$Country[risk_adjusted_age$Country == "GuineaBissau"] <- "Guinea-Bissau"
risk_adjusted_age$Country[risk_adjusted_age$Country == "RepublicofKorea"] <- "SouthKorea"
risk_adjusted_age$Country[risk_adjusted_age$Country == "RepublicofMoldova"] <- "Moldova"
risk_adjusted_age$Country[risk_adjusted_age$Country == "RussianFederation"] <- "Russia"

risk_adjusted_age$Country[risk_adjusted_age$Country == "StateofPalestine"] <- "Palestine"
risk_adjusted_age$Country[risk_adjusted_age$Country == "SyrianArabRepublic"] <- "Syria"
risk_adjusted_age$Country[risk_adjusted_age$Country == "TFYRMacedonia"] <- "Macedonia"
risk_adjusted_age$Country[risk_adjusted_age$Country == "UnitedStatesofAmerica"] <- "UnitedStates"
risk_adjusted_age$Country[risk_adjusted_age$Country == "UnitedStatesVirginIslands"] <- "VirginIslands"
risk_adjusted_age$Country[risk_adjusted_age$Country == "VietNam"] <- "Vietnam"
risk_adjusted_age$Country[risk_adjusted_age$Country == "TimorLeste"] <- "Timor"

simple_model_data <- full_join(simple_model_data, risk_adjusted_age)
```

# Calculate population of each country by comparing total cases to total cases per million.

```{r}

simple_model_data <- simple_model_data[!is.na(simple_model_data$Country),]

population_size <- c()
i = "SouthAfrica"
for(i in unique(simple_model_data$Country)) {
  temp_data <- simple_model_data %>% filter(Country == i)
  
  if(is.na(max(temp_data$total_cases)) |
     max(temp_data$total_cases) == 0 |
     is.na(max(temp_data$total_cases_per_million))) {
    population_size <- c(population_size, rep(0, length(temp_data$total_cases)))
  } else(
    population_size <- c(population_size, rep(max(temp_data$total_cases) / max(temp_data$total_cases_per_million), length(temp_data$total_cases)))
  )
}

sum(is.na(population_size))

simple_model_data <- cbind.data.frame(simple_model_data, population_size)

```

