---
title: "04 - Simple Regression Model"
author: "Murray McKechnie"
date: "14/5/2020"
output: html_document
---

GLM Predicts deaths x days after initial outbreak given average age, GDP (PPP), and Obesity rate.

Other options:

  Total Number of deaths t days after 1st infection.
  Total Number of deaths t days after infections crossed 0.1 per million people.
  
Could use t as a choice before fitting the model, or as an independent variable. In latter a linear model is not ideal, and perhaps a logistic growth model would be better.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
library(ciTools)
library(tidyverse)

simple_model_data <- readRDS("03_Workspace/simple_model_data.Rds")
```

# Select the countries for the models, and obtain their relevant variable values

We only choose countries that have reached a peak in Covid-19 deaths, have population > 3 million, and have experienced more than 50 Covid-19 deaths. We exclude China. They must also have data for all the variables used in the model.

```{r}
model_variables <- simple_model_data %>%
  group_by(Country) %>%
  mutate(average_gov_response = mean(case_when(days_since_deaths_crossed_0.1_per_million >= 0 & date <= "2020-05-09" ~ StringencyIndex,
                                               TRUE ~ NaN), na.rm = T)) %>%
  group_by(Country, Obesity_rate, gdp_estimate_ppp, Median_Age, population_size, peak_deaths_already_occurred_indicator,
           age_adjusted_risk, Density_km, Inequality_spatial_transformed, gov_response_at_0.1_deaths_per_million, average_gov_response) %>%
  mutate(total_deaths = max(total_deaths, na.rm = T)) %>%
  filter(deaths_5_day_averages == max(deaths_5_day_averages, na.rm = T)) %>% # Just keeps the day(s) of the peak for each country
  summarise(total_deaths_per_million = max(total_deaths_per_million, na.rm = T),
            smooth_new_deaths_per_million = max(deaths_5_day_averages, na.rm = T),
            total_deaths = max(total_deaths, na.rm = T)) %>%
  filter(!is.na(Obesity_rate) &
         #!is.na(gdp_estimate_ppp) &
         !is.na(total_deaths_per_million) &
         !is.na(age_adjusted_risk) &
         !(is.na(Density_km)) &
         !is.na(gov_response_at_0.1_deaths_per_million) &
         peak_deaths_already_occurred_indicator == 1 &
         population_size >= 3 &
         total_deaths >= 50 &
         Country != c("China")
         ) %>%
  mutate(combined_density_measure = Density_km * Inequality_spatial_transformed)

write.csv(model_variables, file = "04_Results/condensed_data_for_models.csv")
```

# Model 1: Predict total_deaths_per_million using age_adjusted_risk

```{r}
m1 <- glm(data = model_variables, family = Gamma("log"), formula = total_deaths_per_million ~ age_adjusted_risk)

summary(m1)
```

## Compare fitted values to actual values

```{r}
ggplot(data = model_variables, aes(x = m1$fitted.values, y = total_deaths_per_million)) + 
  geom_point() +
  geom_text(aes(label = Country), size = 3, check_overlap = T, nudge_y = 7) +
  geom_line(aes(y = seq(0, 300, length.out = dim(model_variables)[1]), x = seq(0, 300, length.out = dim(model_variables)[1])), linetype = "dashed") +
  labs(title = "Total_DPM ~ age_adjusted_risk",
       subtitle = "For countries that have observed a peak in deaths and have > 50 deaths in total",
       x = "Fitted Values",
       y = "Actual Values")
```

## Check residuals

```{r}
plot(m1)
```

# Model 2: Predict total_deaths_per_million using age_adjusted_risk and Obesity_rate

```{r}
m2 <- glm(data = model_variables, family = Gamma("log"), formula = total_deaths_per_million ~ Obesity_rate + age_adjusted_risk)

summary(m2)
```

## Compare fitted values to actual values

```{r}
ggplot(data = model_variables, aes(x = m2$fitted.values, y = total_deaths_per_million)) + 
  geom_point() +
  geom_text(aes(label = Country), size = 3, check_overlap = T, nudge_y = 7) +
  geom_line(aes(y = seq(0, 300, length.out = dim(model_variables)[1]), x = seq(0, 300, length.out = dim(model_variables)[1])), linetype = "dashed") +
  labs(title = "Total_DPM ~ Obesity_rate + age_adjusted_risk",
       subtitle = "For countries that have observed a peak in deaths and have > 50 deaths in total",
       x = "Fitted Values",
       y = "Actual Values")
```

## Check residuals

```{r}
plot(m2)
```

# Model 3: Predict total_deaths_per_million using age_adjusted_risk, Obesity_rate, and Inequality_spatial_transformed

```{r}
m3 <- glm(data = model_variables, family = Gamma("log"),
          formula = total_deaths_per_million ~ Obesity_rate + age_adjusted_risk + Inequality_spatial_transformed)

# Worse models
# m3b <- glm(data = model_variables, family = Gamma("log"),
#           formula = total_deaths_per_million ~ Obesity_rate + Inequality_spatial_transformed)
# 
# m3c <- glm(data = model_variables, family = Gamma("log"),
#           formula = total_deaths_per_million ~ Obesity_rate + age_adjusted_risk + combined_density_measure)

summary(m3)
```

## Compare fitted values to actual values

```{r}
ggplot(data = model_variables, aes(x = m3$fitted.values, y = total_deaths_per_million)) + 
  geom_point() +
  geom_text(aes(label = Country), size = 3, check_overlap = T, nudge_y = 7) +
  geom_line(aes(y = seq(0, 300, length.out = dim(model_variables)[1]), x = seq(0, 300, length.out = dim(model_variables)[1])), linetype = "dashed") +
  labs(title = "Total_DPM ~ Obesity_rate + age_adjusted_risk + Inequality_spatial_transformed",
       subtitle = "For countries that have observed a peak in deaths and have > 50 deaths in total",
       x = "Fitted Values",
       y = "Actual Values")
```

## Check residuals

```{r}
plot(m3)
```

# Model 4: Predict total_deaths_per_million using age_adjusted_risk, Obesity_rate, and gov_response_at_0.1_deaths_per_million

```{r}
# Include gov response data
m4 <- glm(data = model_variables, family = Gamma("log"), formula = total_deaths_per_million ~ Obesity_rate + age_adjusted_risk + gov_response_at_0.1_deaths_per_million)

summary(m4)
```

## Compare fitted values to actual values

```{r}
ggplot(data = model_variables, aes(x = m4$fitted.values, y = total_deaths_per_million)) + 
  geom_point() +
  geom_text(aes(label = Country), size = 3, check_overlap = T, nudge_y = 7) +
  geom_line(aes(y = seq(0, 300, length.out = dim(model_variables)[1]), x = seq(0, 300, length.out = dim(model_variables)[1])), linetype = "dashed") +
  labs(title = "Total_DPM ~ Obesity_rate + age_adjusted_risk + gov_response",
       subtitle = "For countries that have observed a peak in deaths and have > 50 deaths in total",
       x = "Fitted Values",
       y = "Actual Values")
```

## Check residuals

```{r}
plot(m4)
```

# Model 5: Predict total_deaths_per_million using age_adjusted_risk, Obesity_rate, Inequality_spatial_transformed, and gov_response_at_0.1_deaths_per_million

```{r}
# Include all of above
m5 <- glm(data = model_variables, family = Gamma("log"),
          formula = total_deaths_per_million ~ Obesity_rate + age_adjusted_risk + gov_response_at_0.1_deaths_per_million + Inequality_spatial_transformed)

# Worse model
# m5b <- glm(data = model_variables, family = Gamma("log"),
#           formula = total_deaths_per_million ~ Obesity_rate + age_adjusted_risk + gdp_estimate_ppp + average_gov_response + Inequality_spatial_transformed)

summary(m5)
```

## Compare fitted values to actual values

```{r}
ggplot(data = model_variables, aes(x = m5$fitted.values, y = total_deaths_per_million)) + 
  geom_point() +
  geom_text(aes(label = Country), size = 3, check_overlap = T, nudge_y = 7) +
  geom_line(aes(y = seq(0, 300, length.out = dim(model_variables)[1]), x = seq(0, 300, length.out = dim(model_variables)[1])), linetype = "dashed") +
  labs(title = "Total_DPM ~ Obesity_rate + age_adjusted_risk + gov_resp + spatial_inequality",
       subtitle = "For countries that have observed a peak in deaths and have > 50 deaths in total",
       x = "Fitted Values",
       y = "Actual Values")

ggsave("04_Results/Best_Fitted_vs_Actual_Values.png")
```

## Predict total_deaths_per_million for countries not included in the model

```{r}
predictions <- add_pi(tb = simple_model_data %>%
                             filter(!(Country %in% model_variables$Country) & date == "2020-05-11") %>%
                             mutate(total_deaths_per_million = 0), fit = m5) %>%
                 select(Country, pred, LPB0.025, UPB0.975) %>%
                 rename(Predicted_total_deaths_at_peak = pred)

write.csv(predictions, "04_Results/best_model_predictions_for_out_of_sample_countries.csv")

# Compare fits of countries included to actual values - table
fits <- cbind.data.frame(model_variables %>%
                           ungroup() %>%
                           select(Country, total_deaths_per_million), m5$fitted.values) %>%
                           rename(Fitted_Values = `m5$fitted.values`,
                                  Actual_values = total_deaths_per_million)

write.csv(fits, "04_Results/best_model_fits_for_in_sample_countries.csv")

# Correlation matrix of variables we included in model 5

as.data.frame(cor(model_variables %>% ungroup() %>% select(Obesity_rate, age_adjusted_risk, gdp_estimate_ppp, gov_response_at_0.1_deaths_per_million, Density_km, Inequality_spatial_transformed)))
```

## Check residuals

```{r}
plot(m5)
```
